%*******************************************************
% Publications
%*******************************************************
% \pdfbookmark[1]{Acknowledgments}{acknowledgments}
% \phantomsection
% \addtocontents{toc}{\protect\vspace{\beforebibskip}}%
% \addcontentsline{toc}{chapter}{\tocEntry{Publications Related to this Thesis}}%
\addchap{About the Author}\label{ch:pubs}
\chaptermark{About the Author}

\begingroup

Pratik R. Gupte was born in India in 1993.
After schooling in Hyderabad and an undergraduate degree in zoology from St. Xavier's College, Mumbai in 2014, he worked on field projects in southern India, Ladakh, and in South Africa.
In 2017 he received a master's degree from the University of Kiel, as part of the International Master's in Applied Ecology, a programme spread over France, Portugal, Germany, and Ecuador.
His master's thesis on families of wintering Arctic geese saw fieldwork in the Russian Arctic and the Netherlands.
In 2018, he began his PhD in Franjo Weissing's lab at the University of Groningen.
Pratik is broadly interested in the spatial ecology of animals, especially birds.
Recognising that the current model of academic science is unsustainable, he left the academic career track, and is now a research software engineer at the London School of Hygiene and Tropical Medicine, where he develops epidemiological models to inform policy responses to disease outbreaks.

\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

% set the separation between entries with different author name first letters
% to zero, and this suppresses letter printing in the margin as in the main bib
\bibinitsep=0

\begin{refsection}
    \small
    \nocite{gupte2021a,gupte2022c,gupte2022d,thaker2019,nathan2022,netz2022,ramesh2022,
        bijleveld2021,rimbach2022un,gupte2019} % is local to to the enclosing refsection
    \printbibliography[title=Author Publications]
\end{refsection}

\begin{refsection}
    \small
    \nocite{gupte2022,gupte2022e,gupte2022b,gupte2021b,gupte2020a,gupte2022f,netz2021b,netz2022a,netz2021b} % is local to to the enclosing refsection
    \printbibliography[title=Data and Code]
\end{refsection}

\endgroup

{ \begin{center} \barfont{-.-} \end{center} }

% \emph{Attention}: This requires a separate run of \texttt{bibtex} for your \texttt{refsection}, \eg, \texttt{ClassicThesis1-blx} for this file. You might also use \texttt{biber} as the backend for \texttt{biblatex}. See also \url{http://tex.stackexchange.com/questions/128196/problem-with-refsection}.
